Bayesian Sparsification Methods for Deep Complex-valued Networks

### 2020-06-22 paper/paper.tex
8 improving the presentation of the findings  * moved expected outcome from intro section to conclusions  * compressed the preliminaries of the experimental methodology  * wrote up the findings of the MNIST-like section better  * condensed CIFAR10 section and drew clearer conclusions  * added some thought stubs to conclusion

% our contribution
Sparse Variational Dropout is a Bayesian Variational Inference method with automatic
parameter relevance determination effect. In this study we extend \emph{Sparse VD} to
$\cplx$VNN, inspired by the results of \citet{gale_state_2019}, and motivated by seldom
application of Bayesian Inference to $\cplx$-valued networks \citep{popa_complex-valued_2018}
and apparent scarcity of compression methods specific to them. We assess the performance-%
compression trade-off of the extension by conducting a large-scale numerical study on
image classification on MNIST-like and CIFAR10 datasets and music transcription on MusicNet.
%
% THIS SHOULD GO TO CONCLUSIONS!
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
We expect the overall conclusion of \citet{gale_state_2019} to carry over to $\cplx$VNN,
since there is nothing fundamentally different in $\cplx$-valued versions of the methods
they study, except for the need to operate on pairs of parameters.

@@

Variational Dropout uses \emph{fully factorized Gaussian} approximation $
  q_\theta(\omega)
    = \prod_j \mathcal{N}(\omega_j \vert\, \mu_j, \alpha_j \mu_j^2)
$ and factorized scale invariant log-uniform prior $\pi(\omega)$ with $
  \pi(\omega_j) \propto \lvert \omega_j \rvert^{-1}
$. \citet{molchanov_variational_2017} noticed that $\alpha_j$ reflects the relevance of the
parameter $\omega_j$ it is associated to by being the ratio of its squared mean to its
effective variance. Based on this observation they proposed \emph{Sparse Variational Dropout},
a modification that enables automatic model sparsification by optimizing $\alpha_j$ for each
individual parameter. \citet{louizos_bayesian_2017} extended the idea to structured sparsity
by considering hierarchical prior and variational approximation. They grouped the parameters
$\omega_j$ and coupled them within each one through a shared latent variable, which on the
whole enabled pruning entire input features in each layer.
% {louizos_bayesian_2017} eq. (11) variance -> gaussian entropy -> informative bits
%

@@

We study the performance-compression trade-off of the proposed Bayesian sparsification
method for $\cplx$-valued networks presented above on image classification and music
annotation tasks.
%
Following \citet{wolter_complex_2018} and \citet{trabelsi_deep_2018}, the class logit
scores are taken as real part of the complex-valued output of a network.
%
% image features
Since image data is not naturally $\cplx$-valued, we preprocess it by either embedding
% $\real \hookrightarrow \cplx$
$\real$-data into $\cplx$ assuming $\Im z = 0$ (\texttt{raw}), or applying the two-dimensional
Fourier Transform (\texttt{fft}), centering the lower frequencies. We do not train an
auxiliary model to synthesize the imaginary component from the real data, proposed by
\citet{trabelsi_deep_2018}.

% setup and common experiment settings
For each dataset we conduct a set of experiments in which we try out all combinations of
model kinds ($\real$ or $\cplx$), sparsification methods (VD or ARD) and input features.
Where reasonable, we also study the effects of halving ($\tfrac12\cplx$) and doubling ($2\real$)
of the size of intermediate layers in each network \citep{monning_evaluation_2018}.
%
Each experiment is replicated \emph{five} times to take into account random initialization,
random order of the mini-batch during SGD, stochastic outputs of the intermediate layers,
and non-determinism of single-precision computations on GPUs.

Each network within every experiment passes through three successive stages of training:
pre-training (\emph{``dense''}), sparsification (\emph{``sparsify''}), and fine-tuning
(\emph{``fine-tune''}) (sec.~\ref{sub:staging}). The parameters and optimizer state are
passed between stages, except for last one, which inherits just the parameters.
%
Every experiment uses ADAM optimizer, resets the learning rate to $10^{-3}$ before each
stage, and applies global $\ell_2$-norm gradient clipping at $0.5$.

% write a note on the incompressible parameters (bias, batch norm etc.)
The compression rate is calculated based on the number of IEEE754 floating point values
needed to store the model and equals $
  \tfrac{n_\mathtt{par}}{n_\mathtt{par} - n_\mathtt{zer}}
$, where $n_\mathtt{zer}$ is the number of explicit zeros, $n_\mathtt{par}$ is the total
number of floats. In particular, each parameter is counted once in a $\real$-valued network,
but twice in a $\cplx$-network. Each model has a compression limit, determined by biases,
shift and scaling in $\real$- and $\cplx$-valued batch normalization layers.

@@

% The training consists of three successive stages: ``dense'' $\to$ ``sparsify'' $\to$ ``fine-tune''.

At the \emph{``dense''} stage every network is fit ``as-is'', retaining its original
non-Bayesian architecture, and using only the likelihood term from \eqref{eq:sgvb_estimator} without
variational dropout.

During the \emph{``sparsify''} stage we make every layer Bayesian and apply $\cplx$-variational
dropout sec.~\ref{ssub:vd_prior} (or sec.~\ref{ssub:ard_prior}) or its $\real$ counterpart.
We inject a coefficient $
  C \in (0, 1]
$ at the KL divergence term in \eqref{eq:sgvb_estimator} to study the attainable sparsity levels:
\begin{equation}  \label{eq:elbo_with_coef}
  - \frac{C}N KL(q_{\theta} \| \pi_{\lambda})
  + \frac1{M} \sum_{k=1}^M
      \log p_{\phi}(z_{i_k} \mid g(\varepsilon_k; \theta))
  \,.
  \tag{\ref{eq:sgvb_estimator}'}
\end{equation}
By varying $C$ we adjust the importance of the sparsifying prior in learning $q_\theta(W)$,
meaning that higher value yields more irrelevant parameters during inference and makes
layers sparser upon termination of this stage.
% RFC: citation needed fot this "trick". Cite $\beta$-VAE maybe?

% on compresssion
Between \emph{``sparsify''} and \emph{``fine-tune''} stages we compute masks of non-zero
weights in each layer based on the relevance scores $\alpha$ (sec.~\ref{sec:variational_dropout}).
Since the $q_\theta$ factorizes into univariate distributions, a $\cplx$ or $\real$ parameter
is considered non-zero if $\log \alpha \leq \tau$ for $
  \alpha = \tfrac{\sigma^2}{\lvert\mu\rvert^2}
$, or zero otherwise.

The \emph{``fine-tune''} stage reverts the network back to non-Bayesian layers, applies
sparsity masks to its parameters and proceeds with training the kept parameters as during
the ``dense'' stage. Unless specified otherwise the network is initialized to the mode
of the learnt $q_\theta(W)$.

% on chosing the threshold
The threshold $\tau$ is picked so, that the remaining non-zero parameters are within
$\delta$ relative tolerance of their mode with high probability under the approximate
posterior. Since $q_\theta$ is fully factorized $\real$- or a proper $\cplx$-Gaussian,
the random variable $
  \tfrac{k \lvert w - \mu \rvert^2}
        {\alpha \lvert \mu \rvert^2}
$ is $\chi^2_k$ distributed with $k=1$ ($\real$) or $2$ ($\cplx$).
%
For a generous tolerance $\delta = 50\%$ all values of $\log \alpha$ below $-2.5$ yield
at least $90\%$ chance of a non-zero $\real$/$\cplx$ parameter. We pick $\tau = -\tfrac12$
to retain parameters sufficiently concentrated around their mode, and encourage higher
sparsity, at the same time being aware that $q_\theta$ is merely an approximation.
In comparison, $\tau = 3$ is commonly used as the threshold
\citep{molchanov_variational_2017,kingma_variational_2015}.

@@

We conduct a moderately sized experiment on MNIST-like datasets of $28\times 28$ greyscale
images to study the performance-compression trade-off of the proposed $\cplx$-variational
dropout: MNIST \citep{lecun_gradient-based_1998}, KMNIST \citep{clanuwat_deep_2018},
EMNIST \citep{cohen_emnist_2017} and Fashion-MNIST \citep{xiao_fashion-mnist_2017}.
%
% dataset size
We deliberately use a {fixed} random subset of $10k$ images from the train split of each
dataset to fit the networks and measure the performance with classification \emph{accuracy}
score on the usual test split.
% Limiting the size was done primarily to test the regularization strength of compression
% and to facilitate quicker experiments.

% model architectures
% (fft) C - R - C / 2
% (raw) R - C - 2 * R
We consider two simple architectures for $\real$ and $\cplx$ networks in this experiment,
which have been chosen for the purpose of illustrating the compression and understanding
the effects of experiment parameters. \emph{TwoLayerDenseModel} is a wide dense ReLU
network $784 \to 4096 \to n_\mathtt{out}$, and \emph{SimpleConvModel} is a ReLU net with
two $2d$ $k_5 s_1$ convolutions with filters $20 \to 50$, two $k_2 s_2$ average pooling steps,
and a dense classifier network $800 \to 500 \to n_\mathtt{out}$. The number of intermediate
feature is as specified in $\real$ and $\cplx$ models, is halved in complex models tagged
$\tfrac12\cplx$, and doubled in real models labelled $2\real$.

% experiment setup
Stages (sec.~\ref{sub:staging}) last for $40$, $75$ and $40$ epochs, respectively, in each
experiment. The sparsification threshold $\tau$ is fixed at $-\tfrac12$, the training batch
size is set to $128$ and the base learning rate ${10}^{-3}$ is reduced after the $10$-th epoch
to ${10}^{-4}$ at every stage.
%
% grid specs
We let $
  C \in \{
    \tfrac32 2^{-\tfrac{k}2} \colon k=2, \cdots, 38
  \}
$ in \eqref{eq:elbo_with_coef}, test VD and ARD methods, try \texttt{raw} or \texttt{fft}
input features, and repeat each experiment $5$ times to get a sample of compression-accuracy
pairs.

@@

% observations
Figures~\ref{fig:mnist-like__trade-off__fft} and~\ref{fig:mnist-like__trade-off__raw}
depict the resulting compression-accuracy trade-off on MNIST for the simple models
described above (other datasets are in appendix~\ref{sec:mnist_like_experiments}).
%
The points represent the trade-off of the compressed networks after fine-tuning, while their
tails illustrate the impact of this stage on the accuracy. Transparent horizontal bands on
each plot represent min-max performance spread of an uncompressed network on the test split.

The overarching pattern in all experiments on MNIST-like datasets is that both ARD and VD
compress similarly and ARD offers marginal advantage in terms of performance after fine-tuning.
Fourier features perform marginally worse than raw image data for each model across all
compression rates, and their trade-off for the dense network has a hat-like shape, contrasting
the raw image features, for which the accuracy stays at one level for moderate rates,
and then abruptly decays in high compression regime.

The impact of ``fine-tune'' is mixed, and depends on the dataset and features, although it
frequently improves accuracy in high compression $\times50+$ (high $C$ in \eqref{eq:elbo_with_coef})
for raw image input. On EMNIST Letters and MNIST datasets in the small compression regime the
dense network exhibits a trough in accuracy before fine-tuning stage.

% conclusions
These experiments demonstrate that $\cplx$ Variational Dropout can offer adequate compression
without much loss in performance for $\cplx$-valued networks. Although ARD performs on par
with VD, unlike VD, it uses a proper prior and has a tractable analytic expression for the
KL-divergence \eqref{eq:emp-bayes-kl-div} in both $\real$ and $\cplx$ cases.
% fft: $2\real$ convolutional model trade-off overall better than $\real$, dense -- on par.
% raw: $\cplx$ models trade-off overall better than $\tfrac12\cplx$.
Wider $\real$ and $\cplx$ networks appear to have more favourable trade-off between
compression and accuracy.
% fft: any-$\real$ dense model compresses better than $\cplx$
However, it seems that due to more intrinsic degrees of freedom, $\real$-valued models
perform and compress better than $\cplx$, even when the former uses the same number of
features as the latter.
% $\cplx$-valued networks do not have as much spare capacity as $2 \real$ due to
%  constraints imposed by multiplication in $\cplx$ numbers

@@

We conduct an experiment on the CIFAR10 classification dataset \citep{krizhevsky_learning_2009}
comprising $32\times 32$ color images of $10$ classes to study the performance-compression
trade-off of a moderately large deep network.
% same experimental setup as in MNIST, describe the VGG16 model and its cplx-valued modification
The experiment is set up similarly to MNIST-like experiment, except that we train $\real$
VGG16 network \citep{simonyan_very_2015} and its $\cplx$ variant, in which we have replaced
$\real$-valued layers with their $\cplx$-valued counterparts. $\real$ and $\cplx$ VGG16 have
the same feature sizes in intermediate layers.

% differences with mnist
Unlike experiment in sec.~\ref{sub:mnist_like_datasets}, we consider the raw features only,
use full training split, and allocate $20$, $40$, and $20$ epochs to each stage. During training
the mini-batch of $128$ samples is randomly augmented by randomly flipping horizontally and
cropping. Random cropping is done by zero-padding each side by four pixels and extracting a
random $32\times 32$ patch from the $40\times 40$ intermediate image.
%
We vary $C$ in \eqref{eq:elbo_with_coef} over $
  \{\tfrac32 2^{-\tfrac{k}2} \colon k=7, \cdots, 15\}
$, compare VD and ARD methods, and measure accuracy on CIFAR10 test split.

% compare compressed cplx-vgg against real-vgg
\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    % \includegraphics[width=\columnwidth]{figure__cifar__trade-off/legacy__augmentedcifar10__raw__-0.5.pdf}
    \includegraphics[width=\columnwidth]{figure__cifar__trade-off/appendix__augmentedcifar10__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy frontier for VGG16 ($\real$ and $\cplx$) on CIFAR10.
  }
  \label{fig:figure__cifar10__trade-off}
\end{figure}

% draw some conclusions
Figure~\ref{fig:figure__cifar10__trade-off} shows that, although ARD method offers slightly
less compression, it makes up for it by marginally better accuracy post fine-tune. VGG16 and
$\cplx$-VGG16 exhibit similar declining compression-accuracy trade-off, but due to higher
capacity $\cplx$-valued network yields higher accuracy \citep{monning_evaluation_2018}. Unlike
MNIST, fine-tuning stage significantly improves the test performance, which, we speculate, might
be due to dataset augmentation and deeper model.
% reword this sentence

@@

We test complex ARD (sec.~\ref{ssub:ard_prior}) and VD (sec.~\ref{ssub:vd_prior}) methods
and vary $
  C \in \{\tfrac14, \tfrac12, \tfrac34, 1\} \cdot 10^{-k}
$ for $k=1, 2, 3$, while keeping $\tau$ at $-\tfrac12$. The performance is measured after
``dense'' stage, before the beginning and upon termination of fine-tuning.
%
Additionally, we test the model of \citet{trabelsi_deep_2018}, in which we purposefully
halve the receptive field of the first convolution from $6$ to $3$ (denoted by suffix $k_3$).
The motivation is to test if the handicap introduced by the forced compression of the most
upstream layer can be alleviated by non-uniform compression, induced by variational dropout.
We test only complex VD in this sub-experiment, since prior results have not demonstrated
significant superiority of one method over another.  % it should've been ARD

@@

% why is there such a dramamtic difference in impact of ``fine-tune''?
We provide the following interpretation of the apparent difference in performance impact
of the ``fine-tune'' stage between less than $50\times$ and higher than $100\times$ compression
regimes in figure~\ref{fig:musicnet__trade-off}, also observed in sec.~\ref{sub:mnist_like_datasets}.
%
% * early stopping kick in at around 10-20 epoch
% * performance peaks at 10-15 epoch and then gradually declines
The value of $C$ in \eqref{eq:elbo_with_coef} is a good proxy for the ranking of the final
compression rate since it directly affects the feedback from sparsifying prior. So, with
low $C$ models get less compressed during $50$ epoch allotted for ``sparsify'' stage, which
appears to be insufficient to move model away from the parameters after the dense stage.
It is reasonable therefore to expect that for undercompressed models the fine-tuning stage
acts essentially as a continuation of the pre-training stage. Since we have observed that
longer training invariably deteriorates the validation performance, the ``fine-tune'' stage
leads to overfitting. Figure~\ref{fig:musicnet__early_stopping}, which shows that the models,
which have been sparsified with $C$ less than $\tfrac1{400}$ and have $50\times$ compression,
need considerably less training epochs before early stopping terminates the process.

@@

>>>>

% our contribution
Sparse Variational Dropout is a Bayesian Variational Inference method with automatic
parameter relevance determination effect. In this study we extend \emph{Sparse VD} to
$\cplx$VNN, inspired by the results of \citet{gale_state_2019}, and motivated by seldom
application of Bayesian Inference to $\cplx$-valued networks \citep{popa_complex-valued_2018}
and apparent scarcity of compression methods specific to them. We assess the performance-%
compression trade-off of the extension by conducting a large-scale numerical study on
image classification on MNIST-like and CIFAR10 datasets and music transcription on MusicNet.

@@

Variational Dropout uses \emph{fully factorized Gaussian} approximation $
  q_\theta(\omega)
    = \prod_j \mathcal{N}(\omega_j \vert\, \mu_j, \alpha_j \mu_j^2)
$ and factorized scale invariant log-uniform prior $\pi(\omega)$ with $
  \pi(\omega_j) \propto \lvert \omega_j \rvert^{-1}
$. \citet{molchanov_variational_2017} noticed that $\alpha_j$ reflects the relevance of the
parameter $\omega_j$ it is associated to by being the ratio of its squared mean to its
effective variance. Based on this observation they proposed \emph{Sparse Variational Dropout},
a modification that enables automatic model sparsification by optimizing $\alpha_j$ for each
individual parameter. \citet{louizos_bayesian_2017} extended the idea to structured sparsity
by considering hierarchical prior and variational approximation. They grouped the parameters
$\omega_j$ and coupled them within each one through a shared latent variable, which on the
whole enabled pruning entire input features in each layer.
% {louizos_bayesian_2017} eq. (11) variance -> gaussian entropy -> informative bits
% They also proposed to estiamte bit precision for quantization based on parameter's
% estimated uncertainty (variance or entropy).
%

@@

To verify the proposed $\cplx$-valued variational sparsification methods presented
above and explore their compression-performance trade-off we carry out a numerical
study of $\cplx$VNN for image classification and music transcription.

% image features
Since image data is not naturally $\cplx$-valued, we preprocess it using the natural
inclusion $\real \hookrightarrow \cplx$ (\texttt{raw}, $\Im z = 0$) or applying the
two-dimensional Fourier Transform (\texttt{fft}), centering the lower frequencies. We
do not train an auxiliary network that synthesizes the imaginary component from $\real$
input data \citep{trabelsi_deep_2018}.
%
Following \citet{wolter_complex_2018} and \citet{trabelsi_deep_2018}, the class logit
scores are taken as real part of the complex-valued output of a network.

% setup and common experiment settings
The networks are trained in three successive stages in every experiment: the \emph{``pre-train''}
stage for pre-training the network, the \emph{``sparsify''} stage to determine parameter
relevance using Variational Dropout, and the \emph{``fine-tune''} to train the pruned
network (sec.~\ref{sub:staging}). Network's parameters are initialized with values from
the previous stage.
%
Networks are trained with ADAM optimizer, with the learning rate reset to $10^{-3}$ before
each stage and global $\ell_2$-norm gradient clipping at $0.5$.
% https://pytorch.org/docs/master/_modules/torch/nn/utils/clip_grad.html#clip_grad_norm_

Each experiment is replicated \emph{five} times to account for random effects from
initialization, stochastic gradient optimization, noisy output from intermediate layers,
and non-determinism of computations on GPU.
% experiments took about three weeks on 4 node GTX1080Ti GPUs in fp32

% write a note on the incompressible parameters (bias, batch norm etc.)
The compression rate is calculated based on the number of floating point values needed
to store the network and equals $
  \tfrac{n_\mathtt{par}}{n_\mathtt{par} - n_\mathtt{zer}}
$, where $n_\mathtt{zer}$ is the number of explicit zeros at the ``fine-tune'' stage
and $n_\mathtt{par}$ is the total number of values. In a $\real$-valued network each
parameter counts as \emph{one} value and as \emph{two} values in a $\cplx$VNN. Each
model has a compression limit, determined by biases, shift and scaling in $\real$-
and $\cplx$-valued batch normalization layers.

@@

% The training consists of three successive stages: ``pre-train'' $\to$ ``sparsify'' $\to$ ``fine-tune''.
At the \emph{``pre-train''} stage every network is fit ``as-is'', retaining its original
deterministic architecture and using only the likelihood term from \eqref{eq:sgvb_estimator}.

During the \emph{``sparsify''} stage we make every layer Bayesian and apply $\cplx$-valued
variational sparsification (sec.~\ref{ssub:vd_prior}, ~\ref{ssub:ard_prior}) or its $\real$
counterpart. We inject a coefficient $
  C \in (0, 1]
$ at the KL divergence term in \eqref{eq:sgvb_estimator}:
\begin{equation}  \label{eq:elbo_with_coef}
  - \frac{C}N KL(q_{\theta} \| \pi_{\lambda})
  + \frac1{M} \sum_{k=1}^M
      \log p_{\phi}(x_{i_k} \mid g(\varepsilon_k; \theta))
  \,.
  \tag{\ref{eq:sgvb_estimator}'}
\end{equation}
In contrast to \cite{molchanov_variational_2017}, who anneal $C$ from zero to one during
training, we use constant $C$ and vary it between runs. This allows us to explore the
compression-performance profile by balancing model's likelihood and posterior's penalty
for diverging form the sparsifying prior in \eqref{eq:elbo_with_coef}. In particular,
higher $C$ implies higher sparsity.
% RFC: citation needed fot this "trick". Cite $\beta$-VAE maybe?

% on compresssion
Between \emph{``sparsify''} and \emph{``fine-tune''} stages we compute masks of non-zero
weights in each layer based on the relevance scores $\alpha$ (sec.~\ref{sub:dropout}).
Since $q_\theta$ factorizes into univariate distributions, a $\cplx$ or $\real$ parameter
is considered non-zero iff $\log \alpha \leq \tau$ for $
  \alpha = \tfrac{\sigma^2}{\lvert \mu \rvert^2}
$.
% on chosing the threshold
The threshold $\tau$ is picked so that the remaining non-zero parameters are within
$\delta$ relative tolerance of their mode with high probability under the approximate
posterior. For a univariate $\real$- or a circularly symmetric $\cplx$-Gaussian random
variable $w$, $
  \tfrac{k \lvert w - \mu \rvert^2}
        {\alpha \lvert \mu \rvert^2}
$ is $\chi^2_k$ distributed with $k=1$ ($\real$) or $2$ ($\cplx$).
%
For a tolerance $\delta = 50\%$ values $\log \alpha$ below $-2.5$ yield at least $90\%$
chance of a non-zero $\real$/$\cplx$ parameter. We pick $\tau = -\tfrac12$ to retain
parameters sufficiently concentrated around their mode and encourage higher sparsity,
at the same time being aware that $q_\theta$ is merely an approximation. In comparison,
$\tau = 3$ is commonly used as the threshold
\citep{molchanov_variational_2017,kingma_variational_2015}.

The \emph{``fine-tune''} stage reverts the network back to deterministic layers,
initializes the parameters to the mode of $q_\theta(\omega)$ and proceeds as the
``pre-train'' stage, except for training only the parameters specified by sparsity
masks.

@@

We conduct a moderately sized experiment on MNIST-like datasets of $28\times 28$ greyscale
images to study the performance-compression trade-off of the proposed $\cplx$-variational
dropout: MNIST \citep{lecun_gradient-based_1998}, KMNIST \citep{clanuwat_deep_2018},
EMNIST \citep{cohen_emnist_2017} and Fashion-MNIST \citep{xiao_fashion-mnist_2017}.
%
% dataset size
We deliberately use a \emph{fixed} random subset of $10k$ images from the train split
of each dataset to fit the networks and measure the performance with classification
\emph{accuracy} score on the usual test split.
% Limiting the size was done primarily to test the regularization strength of compression
% and to facilitate quicker experiments.

% model architectures
% (fft) C - R - C / 2
% (raw) R - C - 2 * R
We consider two simple architectures in this experiment, which have been chosen for
the purpose of illustrating the compression and understanding the effects of experiment
parameters. \emph{TwoLayerDenseModel} is a wide dense ReLU network $
  784 \to 4096 \to n_\mathtt{out}
$, and \emph{SimpleConvModel} is a ReLU net with two $2d$ $k_5 s_1$ convolutions with
filters $20 \to 50$, two $k_2 s_2$ average pooling steps, and a classifier head
$800 \to 500 \to n_\mathtt{out}$.
%
For each dataset we experiment with all combinations of model kinds ($\real$ or $\cplx$)
and sparsification methods (VD or ARD). To take into account potential differences in
capacity we consider halving or doubling the number of features in the intermediate layers
\citep{monning_evaluation_2018}. Halved $\cplx$VNN are tagged $\tfrac12\cplx$, and
doubled $\real$-valued networks are labelled $2\real$. For \texttt{fft} we compare
$\{\real, \cplx, 2\real\}$ and for \texttt{raw} -- $\{\tfrac12\cplx, \real, \cplx\}$.

% experiment setup
Stages (sec.~\ref{sub:staging}) last for $40$, $75$ and $40$ epochs, respectively, in each
experiment. The sparsification threshold $\tau$ is fixed at $-\tfrac12$, the training batch
size is set to $128$ and the base learning rate ${10}^{-3}$ is reduced after the $10$-th epoch
to ${10}^{-4}$ at every stage.
%
% grid specs
We vary $
  C \in \{
    \tfrac32 2^{-\tfrac{k}2} \colon k=2, \cdots, 38
  \}
$ in \eqref{eq:elbo_with_coef} and repeat each experiment $5$ times to get a sample
of compression-accuracy pairs.

@@

% observations
Figures~\ref{fig:mnist-like__trade-off__fft} and~\ref{fig:mnist-like__trade-off__raw}
depict the resulting compression-accuracy trade-off on MNIST for the models described
above (see appendix~\ref{sec:mnist_like_experiments}).
%
Each point represents the trade-off of the compressed network after fine-tuning, while
its tail illustrates the impact of this stage on the performance. Transparent horizontal
bands on each plot represent min-max performance spread of the pre-trained uncompressed
network on the test split.

% observations and conclusions
At the same time, ``fine-tune'' stage almost always improves performance in high compression
regime ($\times50+$ high $C$ in \eqref{eq:elbo_with_coef}), likely due to regularization from
high sparsity.
% On EMNIST Letters and MNIST datasets in the small compression regime the dense network
% exhibits a trough in accuracy before fine-tuning stage.
Fourier features catch up to the raw data in terms of performance at high compression
rates $\times100+$ only for the \emph{TwoLayerDenseModel}.
% spectral feature useful residual information?
% For it the compression-performance profile exhibits a ramp-like shape, in contrast
% to the raw features, for which the accuracy stays level for moderate rates, but then
% abruptly drops at high rates.
%
Comparison of $\real$ and $\cplx$ networks with matching architecture, i.e. same effective
layer widths in $2\real$ vs. $\cplx$ and $\real$ vs. $\tfrac12\cplx$, shows that doubled
$\real$ networks perform and compress better than $\cplx$, due higher intrinsic redundancy
unchecked by $\cplx$-arithmetic constraint.
% Convolutional network compress worse because they have less redundant parameters

% a table for mean comparison against C and compression-performance?
%
% fft: $2\real$ convolutional model trade-off overall better than $\real$, dense -- on par.
% raw: $\cplx$ models trade-off overall better than $\tfrac12\cplx$.
% fft: any-$\real$ dense model compresses better than $\cplx$
% $\cplx$-valued networks do not have as much spare capacity as $2 \real$ due to
%  constraints imposed by multiplication in $\cplx$ numbers

@@

Having verified the $\cplx$ variational sparsification method on MNIST-like datasets
and simple models, we turn to the CIFAR10 dataset comprising $32\times 32$ colour
images of $10$ classes \citep{krizhevsky_learning_2009} and focus on the VGG16 network
\citep{simonyan_very_2015}.
%
% same experimental setup as in MNIST, describe the VGG16 model and its cplx-valued modification
We train the VGG16 network and its $\cplx$ variant, in which we have replaced $\real$-valued
layers with their $\cplx$-valued counterparts. We do not halve or double the features
in any network, since the goal of this experiment is to assess the trade-off for a deep
convolutional network.
%
% differences with mnist
Unlike experiment in sec.~\ref{sub:mnist_like_datasets}, we consider the raw features
only, use full training split and allocate $20$, $40$, and $20$ epochs to each stage.
During training the mini-batch of $128$ samples is augmented by random horizontal flipping
and random cropping, which is done by zero-padding the image with four pixels and extracting
a $32\times 32$ patch from the $40\times 40$ intermediate image.
%
We vary $C$ in \eqref{eq:elbo_with_coef} over $
  \{\tfrac32 2^{-\tfrac{k}2} \colon k=7, \cdots, 15\}
$, compare $\real / \cplx$ VD and ARD methods and measure accuracy on CIFAR10 test split.

% compare compressed cplx-vgg against real-vgg
\begin{figure}[!t]
  \centering
  \begin{subfigure}[b]{1.\columnwidth}  % imcl2019-style swears at this
    % \includegraphics[width=\columnwidth]{figure__cifar__trade-off/legacy__augmentedcifar10__raw__-0.5.pdf}
    \includegraphics[width=\columnwidth]{figure__cifar__trade-off/appendix__augmentedcifar10__raw__-0.5.pdf}
  \end{subfigure}
  \caption{%
    The compression-accuracy profile for the $\real$ and $\cplx$ VGG16.
  }
  \label{fig:figure__cifar10__trade-off}
\end{figure}

% draw some conclusions
The compression-accuracy curve in figure~\ref{fig:figure__cifar10__trade-off} shows that
it is possible to confidently achieve $\approx100\times$ compression of a deep $\cplx$VNN
without losing accuracy, provided the network is fine-tuned after undergoing Variational
sparsification. Regarding methods themselves, $\cplx$-VD and $\cplx$-ARD follow the same
declining compression-accuracy pattern, but for the same setting $C$ the latter provides
slightly less compression with marginally better accuracy.

@@

We test complex ARD (sec.~\ref{ssub:ard_prior}) and VD (sec.~\ref{ssub:vd_prior}) methods
and vary $
  C \in \{\tfrac14, \tfrac12, \tfrac34, 1\} \cdot 10^{-k}
$ for $k=1, 2, 3$, while keeping $\tau$ at $-\tfrac12$. The performance is measured after
``pre-train'' stage, before the beginning and upon termination of fine-tuning.
%
Additionally, we test the model of \citet{trabelsi_deep_2018}, in which we purposefully
halve the receptive field of the first convolution from $6$ to $3$ (denoted by suffix $k_3$).
The motivation is to test if the handicap introduced by the forced compression of the most
upstream layer can be alleviated by non-uniform compression, induced by variational dropout.
We test only complex VD in this sub-experiment, since prior results have not demonstrated
significant superiority of one method over another.  % it should've been ARD

@@

% why is there such a dramamtic difference in impact of ``fine-tune''?
We provide the following interpretation of the apparent difference in performance impact
of the ``fine-tune'' stage between less than $50\times$ and higher than $100\times$ compression
regimes in figure~\ref{fig:musicnet__trade-off}, also observed in sec.~\ref{sub:mnist_like_datasets}.
%
% * early stopping kick in at around 10-20 epoch
% * performance peaks at 10-15 epoch and then gradually declines
The value of $C$ in \eqref{eq:elbo_with_coef} is a good proxy for the ranking of the final
compression rate since it directly affects the feedback from sparsifying prior. So, with
low $C$ models get less compressed during $50$ epoch allotted for ``sparsify'' stage, which
appears to be insufficient to move model away from the parameters after the ``pre-train'' stage.
It is reasonable therefore to expect that for undercompressed models the fine-tuning stage
acts essentially as a continuation of the pre-training stage. Since we have observed that
longer training invariably deteriorates the validation performance, the ``fine-tune'' stage
leads to overfitting. Figure~\ref{fig:musicnet__early_stopping}, which shows that the models,
which have been sparsified with $C$ less than $\tfrac1{400}$ and have $50\times$ compression,
need considerably less training epochs before early stopping terminates the process.

@@

The overarching conclusion from the conducted experiments is that both $\cplx$-ARD and
$\cplx$-VD methods compress similarly to each other, but for the same value of $C$ in
\eqref{eq:elbo_with_coef} ARD yields slightly higher performance post fine-tuning at
the cost of marginally lower compression.

The results of this experiment and previous ones indicate that the $\cplx$ Variational
Dropout offers adequate compression with controllable loss in performance in $\cplx$VNN
and has trade-off profile similar to the $\real$ counterpart. Although ARD performs on
par with VD, unlike it, it uses a proper prior and has a tractable analytic expression
for the KL-divergence \eqref{eq:emp-bayes-kl-div} in both $\real$ and $\cplx$ cases.%
%
% L0 and dropout work in the low-to-mid sparsity range, variational dropout consistently
% ranked behind L0 on Transformer, and was bested by magnitude pruning for sparsity
% levels of 80\% and up. Variational dropout consistently produces models on-par or better
% than magnitude pruning on ResNet-50, and l0 fails to produce sparse models at all.
We expect the overall conclusion of \citet{gale_state_2019} to carry over to $\cplx$VNN,
since there is nothing fundamentally different in $\cplx$-valued versions of the methods
they study, except for the need to operate on pairs of parameters.
